---
title: "Project 1"
author: "Logan"
date: "3/25/2020"
output: html_document
---

```{r}
library(readxl)
housing <- read_excel("Housing.xlsx")
View(housing)
head(housing)
```


#Data Summary
```{r}
sapply(housing, class)
str(housing)
housing.numeric <- housing[, sapply(housing, is.numeric)]
apply(housing.numeric, 2, mean)
apply(housing.numeric, 2, sd)
apply(housing.numeric, 2, range)
```
After looking at the data in various ways above, it seems that there are no missing values and the data does not look to have any typos or strange values. 
The values that require extra explanation are price, size and lot:

*The price of the house seems to be recorded as a dollar amount in the thousands, so a house with a price value of 388 really means $388,000. 

*The size of the house seems to be recorded as square feet in the thousands, so a house with a size value of 2.18 means 2,180 square feet. 

*The lot variable seems to be just the number of acres the lot the house is built on has. 

```{r}
attach(housing)
```

#Exploratory Data Analysis
```{r}
summary(lm(price~size))
```
```{r}
summary(lm(price~lot))
```
```{r}
summary(lm(price~bath))
```
```{r}
summary(lm(price~bedrooms))
```
```{r}
summary(lm(price~agestandardized))
```
```{r}
summary(lm(price~garagesize))
```
From the summaries of the various linear regression models, it can be concluded that lot, bedrooms, and garagesize have a significant effect on the price of a house due to their low p values.

#Initial Modeling 
```{r}
summary(lm(price~lot+bedrooms+garagesize))
```
$$$$














